---
title: "Presentation to the CEO and CFO of Budweiser - DS6306 Project 1"
author: "Todd Garner"
date: "2023-03-04"
output:
  slidy_presentation: 
    highlight: espresso
    fig_height: 25
    fig_width: 17
editor_options:
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(dev = c('pdf', 'png'), 
        fig.align = 'center', fig.height = 50, fig.width = 25, 
        pdf.options(encoding = "ISOLatin9.enc")) 

library(class)
library(caret)
library(e1071)
library(dplyr)
library(jsonlite)
library(ggplot2)
library(ggthemes)
library(tidyverse)
library(gridExtra)
library(knitr)
```

# Presentation to the CEO and CFO of Budweiser Project 1 - DS6306 Week 8 & 9

Pinnacle Economics would like to thank Budweiser for hiring our company to evaluate a series of questions posed by your company. We will first go through the questions sequentially and then wind up with some insights drawn from the statistics derived from the data and some additional insights drawn from other sources.

***Question 1 - 1. How many breweries are present in each state?***

```{r}
#load the "Breweries.cav" data set from the main GitHub repo.  
Brew <- read.csv(file.choose(), header = TRUE)
head(Brew)
tail(Brew)
#isolate this chunk so I don't continually end up having to reload the same data set.  
```

There are 558 individual rows in the Brewery data set, representing what appears to be 558 distinct breweries within the United States' lower 48 states. We will first filter the data set by state to obtain the number of breweries in each state. But, first, lets take a look at the full data set to observe any anomalies. Looking through the data set, there don't appear to be any NAs or empty values.

From the listing we can see that the top 10 states for numbers of breweries is 1. Colorado - 47 2. California 39 3. Michigan - 32 4. Oregon - 29 5. Texas - 28. 6. Pennsylvania - 25 7. Massachusetts - 23 8. Washington - 23 9. Indiana - 22 10. Wisconsin - 20.

```{r}
Brew_state <- Brew %>% count(State, sort = TRUE)
#(Brew_state)

p <-table(Brew_state$State, Brew_state$n)
barplot(Brew_state$n, xlab = "States", ylab = "Frequency", ylim = c(0,50), main = "Number of Breweries by State", names.arg = Brew_state$State, col = "skyblue", axisnames = TRUE, las = 2) 
#Brew_state
#Brew_state$State
```

***Question 2 - Merge beer data with the breweries data. Print the first 6 observations and the last six observations to check the merged file. (RMD only, this does not need to be included in the presentation or the deck.)***

I will first need to load the Breweries data set into RStudio. There are 2,410 rows in this data set. From the top of the data set, there are apparent NA's. These missing values will be addressed shortly.

```{r}
#Load in The "Beers.csv" from the main GitHub repo

Beer <- read.csv(file.choose(), header = TRUE)
head(Beer)
#Beer
```

We'll now merge these two data sets. The "key" in both data sets is the Brewery_ID. They have different column names so one will have to be changed to match the other.

```{r}
names(Brew)[names(Brew)=="Brew_ID"] <- "Brewery_id"
head(Brew)

Beer_Brew <- merge(Beer, Brew, by = "Brewery_id")
#dim(Beer_Brew)
#head(Beer_Brew)
```

***Question - 3 Address the missing values in each column.***

Let's count the NA's for the data set. From the evaluation function below, we can see that there are 62 missing data points for ABV (alcohol by volume) and 1005 IBU (International Bitterness Units) values missing. With 62 out of 2,410 beers for ABV isn't too bad, and there's a possibility of averaging nearby data together, among other options. Missing IBU is a problem as there are approximately half of the data set we'd have to throw out. Not a good idea. It does not appear that from the given data (in other words, extrapolate the IBU from other beers made at the same brewery) we can estimate IBU so we'll have to make do with what we have.

Doing some searching and reading from the Craft Beer nerds (channeling the sentiments of Todd Alstrom, Founder of the Beer Advocate), the common thought is paraphrased as , "It doesn't really matter what the IBU is as it's not a marketing statistic. Consumer's couldn't care less about the IBU. It's used a lot by brewers to compare to other beers. As Chris McClellan wrote on www.craftbeer.com in an article titled,"Last call for IBU's: Fact, Fiction and their Impact on Your beer" He spoke with Steve Gonzalez, senior manager of Small Batch Brewing & Innovation at Stone Brewing Company., *"IBUs are really interesting, but for the most part, we try not to emphasize them too much in anything consumer-facing," said Gonzalez. "It's not really relevant to your enjoyment of the product, and we're constantly hearing about IBUs across the industry being used an important stat when describing beer. Stone uses IBUs as an important quality control too, like most breweries, and while the consumer certainly wants to see it, we're not making new beers to hit a certain IBU threshold."*

So, not being an important statistic (a subjective designation), throwing out 1005 values (including the other more important information) seems like a bad idea. In the categorization of missing values, I think this would come under Missing Data but not relevant to the consumer facing business. Perhaps do a sample average, median, standard deviation and standard error and find the total mean of the values that are there and put those values into the set so that the mean IBU isn't skewed one way or another. To delete half of the data set doesn't make sense. In the end, I decided to insert the median IBU into the NA's so that it doesn't skew the data set but provides for a full data set from which to evaluate the other variables in each observation.

Now, ABV is an important factor both regulatory-wise, marketing-wise and production quality-wise. The range of missing data points leads me to believe that given additional time, the 62 ABV values could be obtained. For example, the Blue Owl Brewery in Austin Texas has 4 beers with missing ABV's in the data set. However, those 4 beers are shown on their company web site along with their ABV and also IBU. For IBU, I'll need to determine how to impute data values for null values given the neighbors. The following chunk will impute the median of the data set members that are present thus not skewing the statistics horribly. As mentioned above, the IBU doesn't appear that valuable to the consumer. Perhaps it might be valuable to the CEO and CFO of Budweiser? Contemplating further, I'm going to impute the values of IBU from the medians of the present IBU and ABV values. This preserves all of the other data points in each observation and doesn't skew the overall metrics of IBU, or ABV for that matter.

```{r}
Beer_Brew$ABV[is.na(Beer_Brew$ABV)] <- median(Beer_Brew$ABV, na.rm=TRUE)
#summary(Beer_Brew)
Beer_Brew$IBU[is.na(Beer_Brew$IBU)] <- median(Beer_Brew$IBU, na.rm = TRUE)
#summary(Beer_Brew)
```

***Question - 4 Compute the median alcohol content and international bitterness unit for each state. Plot a bar chart to compare.***

This seems like a good use of the group_by function and then calculate the median ABV for each group member.

Below, I break this part into two distinct variables: Style (first) and then State (second chunk)

We can see that the bar chart yields interesting results. As there are many styles, it's difficult to glean significantly meaningful conclusions from the median ABV or IBU by style. By state, on the other hand, yields quite visible and productive results. We can see a listing of the highest ABV and IBU but a heat map would yield more illuminating results. That's just ahead.

```{r}
#Compute the median ABV by Style using group_by
ABV_state <- Beer_Brew %>% group_by(Style) %>% summarize(Median_ABV = median(ABV)) 
#head(ABV_state)

#Compute the median IBU by Style using group_by
IBU_state <- Beer_Brew %>% group_by(Style) %>% summarize(Median_IBU = median(IBU)) 
#head(IBU_state)

ABV_IBU <- merge(ABV_state, IBU_state, by = "Style")
#head(ABV_IBU)
#View(ABV_IBU)
#Below is a bar chart of ABV and separately IBU arranged from highest to lowest.  
library(patchwork)
plot1 <- ggplot(data = ABV_IBU, aes(x = reorder(Style, -Median_ABV), y = Median_ABV)) + geom_bar(stat = "identity", fill = "blue") + labs(x = "Style", y = "ABV") + ggtitle("Median Alcohol by Volume by Style") + coord_flip()
plot2 <- ggplot(data = ABV_IBU, aes(x = reorder(Style, -Median_IBU), y = Median_IBU)) + geom_bar(stat = "identity", fill = "red") + labs(x = "Style", y = "IBU") + ggtitle("Median IBU by Style") + coord_flip()
plot1 / plot2

a <- Beer_Brew %>% filter(ABV == max(ABV)) 
print(a)

b <- Beer_Brew %>% filter(IBU == max(IBU)) 
print(b)

```

These are the two data points requested. First, the Highest ABV followed by the Highest IBU separated by Style.

```{r}
#Compute the median ABV by state using group_by
ABV_state <- Beer_Brew %>% group_by(State) %>% summarize(Median_ABV = median(ABV)) 
#head(ABV_state)

#Compute the median IBU by state using group_by
IBU_state <- Beer_Brew %>% group_by(State) %>% summarize(Median_IBU = median(IBU)) 
#head(IBU_state)

ABV_IBU <- merge(ABV_state, IBU_state, by = "State")
#head(ABV_IBU)
#View(ABV_IBU)
#Below is a bar chart of ABV and separately IBU arranged from highest to lowest.  
library(patchwork)
plot1 <- ggplot(data = ABV_IBU, aes(x = reorder(State, -Median_ABV), y = Median_ABV)) + geom_bar(stat = "identity", fill = "blue") + labs(x = "State", y = "ABV") + ggtitle("Median Alcohol by Volume by State")  
plot2 <- ggplot(data = ABV_IBU, aes(x = reorder(State, -Median_IBU), y = Median_IBU)) + geom_bar(stat = "identity", fill = "red") + labs(x = "State", y = "IBU") + ggtitle("Median IBU by State")
plot1 / plot2

a <- Beer_Brew %>% filter(ABV == max(ABV)) 
print(a)

b <- Beer_Brew %>% filter(IBU == max(IBU)) 
print(b)
```

These are the two data points requested. First, the Highest ABV followed by the Highest IBU separated by State.

***Question 5 - Which state has the maximum alcoholic (ABV) beer? Which state has the most bitter (IBU) beer?***

From the Plots generated in the previous step, we can see: Maximum ABV - District of Colombia <insert political joke here> Maximum IBU - West Virginia

***Question 6. Comment on the summary statistics and distribution of the ABV variable.***

```{r}

#Beer_Brew <- data.frame(Beer_Brew)
#summary(Beer_Brew$ABV)
#z <- mean(Beer_Brew$ABV)
#z
#ggplot(ABV_IBU, aes(x = State, y = mean(ABV))) + geom_point()
hist(ABV_IBU$Median_ABV, xlab = "Average Alcohol by Volume", ylab = "Frequency", main = "Summary statistics for Alcohol by Volume across all 50 States") 

t.test(ABV_IBU$Median_ABV, conf.level = .95)
```

From the histogram, we can see that the data is left skewed. Mean is 5.564% alcohol whereas the Median is 5.625% alcohol by volume. Min and Max means are 4.0%/6.25%. Inner Quartile numbers are 5.5% and 5.825%. A relatively wide range of values. DC - Max, Utah - Min.

A heat map might shed some interesting light on the distribution by area and the next chunk shows that the heat map is instructive. We can see some regions where higher alcohol content is prevalent. The western/southwestern states seemed to have higher alcohol content by average. The upper midwest was similar as was the deep south/southeast.

From the t.test we can see that the p-value is incredibly small rejecting the null hypothesis that the mean = zero. The 95% confidence interval is (0.05486303, 0.05716469) with a mean of means of 0.05601386.

```{r}
#In this chunk, I'll wrangle the data into a format whereby I can plot the mean alcohol by state using a heat map and the lower 48 states map.  

#head(ABV_IBU)
#ABV_IBU$State <- tolower(ABV_IBU$State)
#ABV_IBU$State
library(ggplot2)
library(maps)
#Acu = read.csv(file.choose(),header = TRUE) # read in company data
lookup = data.frame(abb = state.abb, State = state.name) #makes a data frame with State name and abbreviation.
#head(lookup)
#lookup$abb <- sort(lookup$abb)
#head(lookup)
#head(ABV_IBU)
ABV_IBU <- ABV_IBU[-4]
#head(ABV_IBU)
names(ABV_IBU)[names(ABV_IBU) == "State"] <- "abb"
#head(ABV_IBU)
#summary(lookup)
#summary(ABV_IBU)
#ABV_IBU
#lookup
#Data set lookup only has 50 states and does not include DC as ABV_IBU does.  So, we must add that and arrange it in the correct order.  
lookup[nrow(lookup) +1,] = c("DC", "DistofColum")
#lookup
lookup <- arrange(lookup, abb, .by_group = FALSE)
#lookup

ABV_IBU_new = cbind(ABV_IBU,lookup)# make one dataset with state names and abb
#head(ABV_IBU_new)
ABV_IBU_new <- ABV_IBU_new[-1]
#head(ABV_IBU_new)
names(ABV_IBU)[names(ABV_IBU) == "abb.1" ] <- "State"
#head(ABV_IBU)

ABV_IBU_new$State <- tolower(ABV_IBU_new$State)
#head(ABV_IBU_new)
names(ABV_IBU_new)[names(ABV_IBU_new) == "State"] <- "region"
#View(ABV_IBU_new)

Median <- ABV_IBU$Median_ABV

#Bring in the states data set that contains lat/lon for constructing the map of the US. 
states <- map_data("state")
#head(states)
map.df <- inner_join(states,ABV_IBU_new, match = all)
#View(map.df)

#I made a data set of the locations of the Budweiser breweries and worked relentlessly to superimpose those on the heat map to no avail.  This was a crushing defeat. 
bud_sites <- read.csv(file.choose(), header = TRUE)
#View(bud_sites)


map.df <- map.df[order(map.df$order),]
ggplot(map.df, aes(x=long,y=lat,group=group))+
geom_polygon(aes(fill=Median_ABV))+
geom_path()+
scale_fill_gradientn(colours=rev(heat.colors(10)),na.value="grey90")+ggtitle("Median Alcohol By Volume by State")
coord_map()
```

***Question 7. Is there an apparent relationship between the bitterness of the beer and its alcoholic content? Draw a scatter plot. Make your best judgment of a relationship and EXPLAIN your answer.***

Let's start with the scatter plot. The scatter plot of Median ABV against Median IBU for each state with a linear regression line plot through the data shows a relatively strong positive relationship between ABV and IBU. So, the answer is a resounding yes! There is a relationship between ABV and IBU. A positive correlation, meaning the higher the IBU, the higher the ABV. From my own personal experience while living in Washington State, microbrews are much more prevalent than in states such as Texas, but that is changing. The heavier the taste (IPA's specifically), the higher the alcohol content.

```{r}
library(tidyverse)
ggplot(ABV_IBU, aes(x = Median_ABV, y = Median_IBU)) +
  theme_classic() +
  geom_point(
    mapping = aes(Median_ABV, Median_IBU)) + xlab("Median ABV") + ylab("Median IBU") +
    ggtitle("Median ABV versus Median IBU for each state (the dots) with a blue linear regression line plotted for all of the state data points") +
    geom_smooth(method="lm")
```

***Question 8. Budweiser would also like to investigate the difference with respect to IBU and ABV between IPAs (India Pale Ales) and other types of Ale (any beer with "Ale" in its name other than IPA). You decide to use KNN classification to investigate this relationship. Provide statistical evidence one way or the other. You can of course assume your audience is comfortable with percentages ... KNN is very easy to understand conceptually. In addition, while you have decided to use KNN to investigate this relationship (KNN is required) you may also feel free to supplement your response to this question with any other methods or techniques you have learned. Creativity and alternative solutions are always encouraged.***

First, let's extract the character pattern, "India Pale Ale" or "IPA". Looking over the data set, there are many in between (gray area) names like "American Pale Ale." I'd like to create a data.frame that contains just these patterns. There is also the Name.x of the beer and then another column called "Style". The data.frame created above "Beer_Brew" contains all of the data joined by the key Brewery_ID. This seems like a good place to start.

```{r}
# Create KNN Train and Test Datasets ####

train = data.frame(Beer_Brew) # Full data set with all beers and breweries with NA's replaced with proper median values
#View(train)
#dim(train)
#str(train)

train2 = train[grepl("Ale|IPA", train$Style), ] # Everything Ale and IPA

train2$Classification = ifelse(grepl("\\bAle\\b", train2$Style), "Ale", "IPA" ) # Change all "Styles" to either "Ale" or "IPA"

# ------------------------ #

train3 = train2[!grepl("\\bIPA\\b", train2$Style), ] # ONLY ALE 

test2 = train2[grepl("\\bIPA\\b", train2$Style), ] # ONLY IPA

#str(train)
#train
#str(train3)

#str(test2)
```

Below, I'll use KNN to evaluate the relationship between ABV and IBU and specifically the difference for "Ale" and "IPA".

```{r}


#dim(Beer_Brew)

test2 = train2[grepl("\\bIPA\\b", train2$Style), ] # ONLY IPA
#dim(test2)

IPA_ALE <- Beer_Brew[grepl("Ale|IPA", train$Style), ] # Both Ale and IPA
IPA_true_test <- Beer_Brew[grepl("\\bIPA\\b", train$Style), ] # Only IPA
IPA_ALE$classification <- ifelse(grepl("\\bAle\\b", IPA_ALE$Style), "Ale", "IPA" ) #Change all to either Ale or IPA


#I will use the Ale data set to train the model.  I will perform a 75/25 split between training and testing sets, using just the Ale data set.  
set.seed(6)
splitPerc = .75

trainIndices = sample(1:dim(IPA_ALE)[1],round(splitPerc * dim(IPA_ALE)[1]))
train = (IPA_ALE[trainIndices,])
test = (IPA_ALE[-trainIndices,])

#Perhaps we should try to scale ABV & IBU as their relative number values are very dissimilar.  
train_scale <- scale(train[, 4:5])
test_scale <- scale(test[, 4:5])
IPA_scale <- scale(IPA_ALE[,4:5])

IPA_ALE %>% ggplot(aes(x = ABV,IBU,color = classification)) + geom_point() + ggtitle("Relationship between ABV & IBU and distribution of IPA and Ale's")
# k = 5
classifications = knn(train_scale[,c(1:2)],IPA_scale[,c(1:2)],IPA_ALE$classification[1:length(train_scale[ ,1])], k = 5)

cm <- table(classifications, IPA_ALE$classification)
confusionMatrix(table(classifications, IPA_ALE$classification))
```

From the data results using the confusion matrix, we can see there the accuracy is just above 50% (ranging from 54% to 57%) favoring Ale. To me, this suggests that if randomly selecting a beer out of a selection of 2,410 available unique beers, that there is a slight positive chance of selecting an Ale. It also suggests that given the ABV & IBU, there is not much difference between IPA and Ale. For the CEO and CFO of Budweiser, it makes sense to test the market with an Ale or IPA segmented Budweiser product with a slight favor of creating an Ale rather than an IPA. At least from the outset, this suggests that a Budweiser Ale product might have a slightly better chance of success against an IPA styled beer.

Question 9 - Find one other useful inference from the data that you feel Budweiser may be able to find value in. You must convince them why it is important and back up your conviction with appropriate statistical evidence.

InBev is a massive conglomerate and the largest beer brewer in the world. It has every conceivable type of beer, including IPA across the globe. But, let's focus on Budweiser and the North American market. Budweiser has partnerships in other craft beers around the country and in Canada as well as Mexico. Most of these are lagers, but heavier than Budweiser beers in IBU and ABV.

My research indicates that the 4% to 6% ABV beers are losing market share to the higher (and lower) ABV markets. It makes sense to explore moving into these areas as these higher alcohol content beers also garner a higher price tag. Premium beer, premium price!

But wait, Budweiser has been around since the early 1800's. It's spent billions upon billions of dollars marketing the Bud name and it's offshoots, Bud Light, Bud Extra, etc. Does it make sense to dilute the beer of the common Joe "six-pack"?

That market is strong, but the craft beers are gaining ground. It makes sense to grab onto some of that market as well. From the Beeradvocate.com (admittedly an online community made up of users that have been "PUTTING THE BEER INTO GEEKS SINCE 1996"), their top 250 beers is a real time list made up of rankings and weightings using "Bayesian" methods to construct the top favorites among beer "nerds." Here is what surveys tell us.

#Load in the "Beer_types.csv" from the main GitHub repo

```{r}
USA_brew <- read.csv(file.choose(), header = TRUE)
head(USA_brew, 10)
#head(USA_brew)
#write.csv(type_number, "Beer_types.csv", row.names = T)

```

These are the top Styles of beer voted into the Beer Advocate's Top 250. This is your target market. Higher alcohol, higher premium prices. Bud had already started "partnering" with brewers all over the country. Continue this march by adding some of these "higher octane" brands.

```{r}
#Load in the csv titled "2021-State-Breakdown.csv" from the main GitHub repo to see the Economic Impact craft beers make on a state's economy.  


econ_impact <- read.csv(file.choose(), header = TRUE)
eco <- head(econ_impact,10)
eco

```

These are the areas where the penetration has already begun. These microbrews have done the heavy lifting. With Budweiser's ownership, you provide capital, distribution, process refinement, economies of scale and clout. Let them keep their brands and leadership, with some guardrails. This is a way that Budweiser can quietly pursue this market without disrupting your key demographic, Joe "six-pack". And, you can claim some of the newer market dollars that are gaining momentum. Just add water and you can enjoy the spoils that these microbreweries are starting to capture.

Extra: ChatGPT - When asked the following question, the answers were provided below:

Question: ***"Please explain America's preference for craft beers like India Pale Ales and other Ales versus Budweiser and Coors"***

Answer: Craft beer has become increasingly popular in the United States over the past few decades, and there are several reasons why people may prefer craft beers like IPAs and ales over mass-produced beers like Budweiser and Coors.

Flavor: Craft beers tend to have more complex and unique flavors than mass-produced beers. Many craft breweries experiment with different ingredients and brewing techniques to create new and interesting flavor profiles that beer enthusiasts appreciate. IPAs and ales, in particular, are known for their bold and hoppy flavors, which can be a refreshing change from the lighter, blander taste of mass-produced beers.

Variety: Craft breweries often offer a wider variety of beer styles than mass-produced breweries. In addition to IPAs and ales, craft breweries may offer stouts, porters, Belgian-style beers, and more. This variety appeals to beer enthusiasts who enjoy trying new and different beers.

Local support: Many people like to support local businesses, and craft breweries are often seen as part of the local community. By drinking craft beer, consumers feel like they are supporting their local economy and contributing to the growth of small businesses.

Authenticity: Craft breweries are often seen as more authentic and genuine than mass-produced breweries. Craft breweries are typically owned and operated by people who are passionate about beer and brewing, and who put a lot of care and attention into each batch. This authenticity resonates with consumers who value craftsmanship and quality.

Health: Some people may perceive craft beer as healthier than mass-produced beers. Craft breweries often use all-natural ingredients and avoid additives and preservatives, which can be seen as healthier and more wholesome.

In summary, the preference for craft beers like IPAs and ales over mass-produced beers like Budweiser and Coors can be attributed to factors such as flavor, variety, local support, authenticity, and health.

When ChatGPT was asked a follow on question, here is the answer provided:

Question: ***"Please explain a brewers economic preference for craft beers like India Pale Ales and other Ales versus Budweiser and Coors"***

Answer: From an economic perspective, craft breweries may have a preference for brewing IPAs and other ales over mass-produced beers like Budweiser and Coors for several reasons:

Higher margins: Craft beers typically command higher prices than mass-produced beers. This is because craft breweries often use higher-quality ingredients and brewing techniques, which can result in a more expensive product. Additionally, consumers are often willing to pay a premium for the unique flavors and variety offered by craft beers. As a result, craft breweries can often earn higher profit margins on their beer sales.

Niche market: While mass-produced beers have broad appeal and target a wide demographic, craft beers appeal to a more niche market of beer enthusiasts who are willing to pay a premium for unique and high-quality beer. By targeting this niche market, craft breweries can differentiate themselves from mass-produced breweries and avoid competing solely on price.

Brand differentiation: By focusing on IPAs and ales, craft breweries can differentiate their brand from mass-produced breweries. While Budweiser and Coors are known for their light, refreshing lagers, craft breweries can differentiate themselves by offering a wider variety of beer styles and flavors. This can help craft breweries build a loyal following among beer enthusiasts who value experimentation and creativity.

Local support: Craft breweries often have a strong local following, which can help them to build a loyal customer base and support their growth. By focusing on IPAs and ales, which are popular among beer enthusiasts, craft breweries can attract a local following of consumers who appreciate the unique flavors and variety offered by craft beer.

In summary, craft breweries may have an economic preference for brewing IPAs and other ales over mass-produced beers like Budweiser and Coors because they can earn higher profit margins, target a niche market of beer enthusiasts, differentiate their brand, and build a strong local following.
